# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r loadData , echo=TRUE,cache=TRUE}
data <- read.csv(unz("activity.zip", "activity.csv"), header=TRUE,colClasses = c("numeric","Date","integer"),sep=",")

```


## What is mean total number of steps taken per day?
```{r centralValues , echo=TRUE,cache=TRUE,results='markup'}
dailySteps = with(data,tapply(X = steps,INDEX = as.factor(date),FUN = sum,na.rm = TRUE))
hist(dailySteps,col = "green")
dev.copy(png,file = 'figure/plot1.png',width = 480,height = 480,units = 'px')
dev.off
dailyMeanSteps = with(data,tapply(X = steps,INDEX = as.factor(date),FUN = mean,na.rm = TRUE))
dailyMedianSteps = with(data,tapply(X = steps,INDEX = as.factor(date),FUN = median,na.rm = TRUE))
as.data.frame(dailyMeanSteps)
as.data.frame(dailyMedianSteps)
```

## What is the average daily activity pattern?
```{r dailyActivityPattern , echo=TRUE,cache=TRUE,results='markup'}
averageStepsPerFiveMinutes <-with(data,tapply(X = steps,INDEX = as.factor(interval),FUN = mean,na.rm = TRUE))
par(bg = 'white')
plot(row.names(averageStepsPerFiveMinutes),as.vector(averageStepsPerFiveMinutes),type = 'l',xlab = "Interval",ylab = "Average Steps",col = 'red')
dev.copy(png,file = 'figure/plot2.png',width = 480,height = 480,units = 'px')
dev.off
maxVal <- max(averageStepsPerFiveMinutes)
maxValueInterval <- names(averageStepsPerFiveMinutes)[which(as.vector(averageStepsPerFiveMinutes) == max(averageStepsPerFiveMinutes))]
```

The average number of steps (averaged across all the days) has the maximum value of `r maxVal`. This value is attained during the interval starting at `r maxValueInterval` minutes 



## Imputing missing values
```{r imputingMissing ,echo=TRUE,cache=TRUE,results='markup'} 
numMissingValues <- sum(is.na(data$steps))
meanDailyActivity <- data.frame(interval = as.numeric(names(averageStepsPerFiveMinutes)),meanSteps = as.vector(averageStepsPerFiveMinutes))
mergedData <- merge(x = data,y = meanDailyActivity,by.x = "interval",by.y  = "interval")
mergedData$steps[which(is.na(mergedData$steps))] <- mergedData$meanSteps[which(is.na(mergedData$steps))]
dailyStepsNew = with(mergedData,tapply(X = steps,INDEX = as.factor(date),FUN = sum,na.rm = TRUE))
hist(dailyStepsNew,col = "blue")
dev.copy(png,file = 'figure/plot3.png',width = 480,height = 480,units = 'px')
dev.off
dailyMeanStepsNew = with(mergedData,tapply(X = steps,INDEX = as.factor(date),FUN = mean,na.rm = TRUE))
dailyMedianStepsNew = with(mergedData,tapply(X = steps,INDEX = as.factor(date),FUN = median,na.rm = TRUE))
as.data.frame(dailyMeanStepsNew)
as.data.frame(dailyMedianStepsNew)

```

Number of rows with missing values in the original Data  = `r numMissingValues`
Difference between two Datasets :The new data is more like normal distribution in comparison to one with missing values

## Are there differences in activity patterns between weekdays and weekends?
```{r weekdayAnalysis ,echo=TRUE,cache=TRUE,results='markup'} 
mergedData$dayType <-  ifelse(weekdays(mergedData$date)== "Sunday" | weekdays(mergedData$date)== "Saturday","weekend","weekday")
weekdayData <- mergedData[mergedData$dayType == "weekday",]
weekendData <- mergedData[mergedData$dayType == "weekend",]
averageStepsWeekday <-with(weekdayData,tapply(X = steps,INDEX = as.factor(interval),FUN = mean,na.rm = TRUE))
averageStepsWeekend <-with(weekendData,tapply(X = steps,INDEX = as.factor(interval),FUN = mean,na.rm = TRUE))
par(mfrow=c(2,1),bg = 'white')

plot(row.names(averageStepsWeekday),as.vector(averageStepsWeekday),type = 'l',xlab = "Interval",ylab = "Average Steps",col = 'blue',main = "Weekday")
plot(row.names(averageStepsWeekend),as.vector(averageStepsWeekend),type = 'l',xlab = "Interval",ylab = "Average Steps",col = 'blue',main = "Weekend")
dev.copy(png,file = 'figure/plot4.png',width = 480,height = 480,units = 'px')
dev.off
```


